---
title: "Exploratory Analysis"
author: "Mathews Philip Venattu"
date: "30/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(corrplot)

bowShock<-read.csv('Data/bow_shock_only_Data.csv')
finalData<-read.csv('Data/15minutes_interval_boeshockData.csv')
```

## Formatting
<br />
Trying to filer only the important predictors from the otiginal dataset


```{r}

bowShock<-bowShock%>%select(c('X_KSM.km.','Y_KSM.km.','Z_KSM.km.','BTotal.nT.','BX_KSM.nT.','BY_KSM.nT.','BZ_KSM.nT.','zcrosslist','ycrosslist','TimeStamp','type_cross','dirn_cross','xcrosslist'))
finalData<-finalData%>%select(c('X_KSM.km.','Y_KSM.km.','Z_KSM.km.','BTotal.nT.','BX_KSM.nT.','BY_KSM.nT.','BZ_KSM.nT.','zcrosslist','ycrosslist','TimeStamp','type_cross','dirn_cross','xcrosslist'))

head(finalData)
head(bowShock)

```

## Outlier Checks

Box plot of Total Magnetic field points <br />

```{r }
ggplot(finalData, aes(x=type_cross , y=BTotal.nT.)) + 
  geom_boxplot(notch=TRUE)
```

Box plot of Magnetic field across X direction <br />

```{r }
ggplot(finalData, aes(x=type_cross , y=BX_KSM.nT.)) + 
  geom_boxplot(notch=TRUE)
```

Box plot of Magnetic field across Y direction <br />

```{r }
ggplot(finalData, aes(x=type_cross , y=BY_KSM.nT.)) + 
  geom_boxplot(notch=TRUE)
```

Box plot of Magnetic field across Z direction <br />

```{r }
ggplot(finalData, aes(x=type_cross , y=BZ_KSM.nT.)) + 
  geom_boxplot(notch=TRUE)
```


## Data Imbalance 

<br />

From the plot we it is clear that BS events occured in a 16 minute time is very low.
```{r }
ggplot(finalData, aes(x=type_cross)) + 
  geom_bar()
```


## Correlation 

<br />

From the plot we it is clear that BS events occured in a 16 minute time is very low.
```{r }
ggplot(finalData, aes(x=type_cross)) + 
  geom_bar()
```

