---
title: "Task 1"
author: "Mathews Philip Venattu and Avantika Singh"
date: "09/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(dplyr)
library(ggplot2)
```

##Reading Data
Using read RDS for getting the data from a rds file format. Created a copy of original variable<br />
```{r}
data1<-readRDS('../Data/Full_Cassini_Master_MP_BS_CMJ_revised2005.rds', refhook = NULL)
data2<-readRDS('../Data/Cass_data2005.rds', refhook = NULL)
copy_data1<-data1
copy_data2<-data2
```
```{r}
head(data1)
```
```{r}
head(data2)
```
## Formatting Date in Data1
Converted doy_cross variable to 2005 Date.<br />

```{r }
copy_data1$date<-as.Date(copy_data1$doy_cross,origin="2005-01-01")
head(copy_data1)
copy_data1$time<-paste(copy_data1$hour_cross,copy_data1$minute_cross,"00",sep=":")
head(copy_data1)
```
## Formatting Date in Data2
Separating the Timestamp into Hour,Minute and converting timestamp into a Date Format.<br /> 
```{r }
copy_data2$TimeStamp<-as.POSIXct(copy_data2$Timestamp.UTC., format="%d/%m/%Y %H:%M")
copy_data2$date<-as.Date(format(copy_data2$TimeStamp, "%Y-%m-%d" ))
copy_data2$hour_cross<-as.integer(format(copy_data2$TimeStamp, "%H" )) 
copy_data2$minute_cross<-as.integer(format(copy_data2$TimeStamp, "%M" )) 
#copy_data2<-copy_data2%>%select(-c(TimeStamp))
head(copy_data2)
```
## Merging Data1 and Data2 by using time variables
Merging two dataset by using time variables.<br /> 
```{r }
leftJoinDf <- left_join(copy_data2,copy_data1,by=c('date','hour_cross','minute_cross'))
head(leftJoinDf)
```
## Filtering all points in MAG 2005 between day 72 and 74
Filtering the points and finding the bowshock and magnetic pause x coordinates.<br /> 
```{r }
d1<-as.Date(72,origin="2005-01-01")
d2<-as.Date(73,origin="2005-01-01")

temp<-filter(leftJoinDf,between(date,d1,d2))
bowShock<-filter(temp,type_cross=="BS")$TimeStamp
MagP<-filter(temp,type_cross=="MP")$TimeStamp
```
## Plotting Graphs
```{r }
ggplot(data=temp, aes(x=TimeStamp , y=abs(BTotal.nT.))) +
  geom_line()+
  ylab("|B|(nT)")+
  geom_vline(xintercept=bowShock, linetype="dotted",colour="blue",size=1.3)+
  geom_vline(xintercept=MagP, linetype="dotted",colour="red",size=1.3)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```
```{r }
ggplot(data=temp, aes(x=TimeStamp)) +
  geom_line(aes(y = BX_KRTP.nT.), color = "red")+
  geom_line(aes(y = BY_KRTP.nT.), color = "green")+
  geom_line(aes(y = BZ_KRTP.nT.), color = "blue")+
  ylab("Bxyz(nT)")+

  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
 
  

```
## Filtering all points in MAG 2005 between day 136 and 138
Filtering the points and finding the bowshock and magnetic pause x coordinates.<br /> 
```{r }
d1<-as.Date(136,origin="2005-01-01")
d2<-as.Date(138,origin="2005-01-01")

temp<-filter(leftJoinDf,between(date,d1,d2))
bowShock<-filter(temp,type_cross=="BS")$TimeStamp
MagP<-filter(temp,type_cross=="MP")$TimeStamp

```
## Plotting Graphs
```{r }
ggplot(data=temp, aes(x=TimeStamp , y=abs(BTotal.nT.))) +
  geom_line()+
  ylab("|B|(nT)")+
  geom_vline(xintercept=bowShock, linetype="dotted",colour="blue",size=1.3)+
  geom_vline(xintercept=MagP, linetype="dotted",colour="red",size=1.3)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```
```{r }
ggplot(data=temp, aes(x=TimeStamp)) +
  geom_line(aes(y = BX_KRTP.nT.), color = "red")+
  geom_line(aes(y = BY_KRTP.nT.), color = "green")+
  geom_line(aes(y = BZ_KRTP.nT.), color = "blue")+
  ylab("Bxyz(nT)")+

  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
 
  

```