---
title: "Logistic Regression"
author: "Mathews Philip Venattu"
date: "24/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(ggplot2)
library(dplyr)
library(caret)
library(ranger)
library(matrixStats)
library(MASS)
```

## Logistic Regresssion Model

Creating indexes for dataset and removing the datapoints that contains `DG` and `SG`

```{r }

data_wider<-readRDS('input_dataset.rds', refhook = NULL)
copy_data<-data_wider
copy_data<-dplyr::select(data_wider, !ends_with("15"))
t<-copy_data[!(copy_data$type_cross=="DG" | copy_data$type_cross=="SC"),]
copy_data<-t
copy_data$type_cross<-factor(copy_data$type_cross)
copy_data$id<-seq.int(nrow(copy_data))
```


## Making the data Balanced

Making the data balanced and creating training and test datset

```{r}

cassini.train_unscaled<-copy_data%>%filter(type_cross=='NE')%>%sample_n(100)%>%rbind(copy_data%>%filter(type_cross=='BS')%>%sample_n(100))%>%rbind(copy_data%>%filter(type_cross=='MP')%>%sample_n(100))
cassini.train_unscaled$type_cross<-factor(cassini.train_unscaled$type_cross)
cassini.test_unscaled<-copy_data[-cassini.train_unscaled$id,]
cassini.test_unscaled<-dplyr::select(cassini.test_unscaled,-id)
cassini.train_unscaled<-dplyr::select(cassini.train_unscaled,-id)
```


## Shuffling Dataset

```{r}
set.seed(42)
rows <- sample(nrow(cassini.train_unscaled))
temp<- cassini.train_unscaled[rows, ]
cassini.train_unscaled<-temp
```


## Scaling the data

```{r}
cassini.train_scaled<-cassini.train_unscaled%>%mutate_if(is.numeric,scale)
means <- lapply(dplyr::select(cassini.train_unscaled,-c(type_cross,dirn_cross)), mean)
sds <- lapply(dplyr::select(cassini.train_unscaled,-c(type_cross,dirn_cross)), sd)
means <- matrix(unlist(means), ncol = 210, byrow = FALSE)
sds <- matrix(unlist(sds), ncol = 210, byrow = FALSE)
```


## Normalizing the TEST DATASET


```{r}

cassini.test_scaled<-dplyr::select(cassini.test_unscaled,-c(type_cross,dirn_cross))

for(col in c(1:dim(cassini.test_scaled)[2])){
  temp<-(cassini.test_scaled[,col]-means[,col])/sds[,col]
  cassini.test_scaled[,col]<-as.matrix(temp)
  temp<-NULL
  
}
cassini.test_scaled$type_cross<-cassini.test_unscaled$type_cross
cassini.test_scaled$dirn_cross<-cassini.test_unscaled$dirn_cross
```

## Logistic Regression Model 1

### BS as 1 and all other events as 0

Removing dirn_cross and creating new variable `new_trainData`.

```{r}
new_trainData<-dplyr::select(cassini.train_scaled,-dirn_cross)
new_testData<-dplyr::select(cassini.test_scaled,-c(dirn_cross))


```

#### Training Data

```{r}
cassini.logTrain<-new_trainData%>%mutate(event_occured=ifelse(type_cross=='NE'|type_cross=='MP',0,1))
cassini.logTrain$event_occured<-as.factor(cassini.logTrain$event_occured)
cassini.logTrain <- dplyr::select(cassini.logTrain,-c(type_cross))

```

#### Test Data

```{r}
cassini.logTest<-new_testData%>%mutate(event_occured=ifelse(type_cross=='NE'|type_cross=='MP',0,1))
cassini.logTest$event_occured<-as.factor(cassini.logTest$event_occured)
cassini.logTest <- dplyr::select(cassini.logTest,-c(type_cross))
```


#### Model 1 Training and Confusion Matrix
```{r}
LOGModel1<- glm(event_occured ~ ., family = "binomial", data = cassini.logTrain)

summary(LOGModel1)

cassini.logTest$prob=predict(LOGModel1,cassini.logTest,type="response")
cassini.logTest$pred <- factor(ifelse(cassini.logTest$prob < .8, 0,1))
confusionMatrix(cassini.logTest$pred, cassini.logTest$event_occured,positive='1')

```


## Logistic Regression Model 2

### MP as 1 and all other events as 0

Removing dirn_cross and creating new variable `new_trainData`.

```{r}
new_trainData<-dplyr::select(cassini.train_scaled,-dirn_cross)
new_testData<-dplyr::select(cassini.test_scaled,-c(dirn_cross))


```

#### Training Data

```{r}
cassini.logTrain<-new_trainData%>%mutate(event_occured=ifelse(type_cross=='NE'|type_cross=='BS',0,1))
cassini.logTrain$event_occured<-as.factor(cassini.logTrain$event_occured)
cassini.logTrain <- dplyr::select(cassini.logTrain,-c(type_cross))

```

#### Test Data

```{r}
cassini.logTest<-new_testData%>%mutate(event_occured=ifelse(type_cross=='NE'|type_cross=='BS',0,1))
cassini.logTest$event_occured<-as.factor(cassini.logTest$event_occured)
cassini.logTest <- dplyr::select(cassini.logTest,-c(type_cross))

```


#### Model 2 Training and Confusion Matrix
```{r}
LOGModel1<- glm(event_occured ~ ., family = "binomial", data = cassini.logTrain)

summary(LOGModel1)

cassini.logTest$prob=predict(LOGModel1,cassini.logTest,type="response")
cassini.logTest$pred <- factor(ifelse(cassini.logTest$prob < .8, 0,1))
confusionMatrix(cassini.logTest$pred, cassini.logTest$event_occured,positive='1')

```
