---
title: "Exploratory Data Analysis"
author: "Mathews Philip Venattu"
date: "31/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(kableExtra)
```


##Reading Data
Using read RDS for getting the data from a rds file format. Created a copy of original variable<br />
```{r}
data_Wider<-readRDS('../Full_Data_Wider.rds', refhook = NULL)
bs_data<-read.csv('../Data/bow_shock_only_Data.csv')
short_Data<-readRDS('../Average_Data.rds', refhook = NULL)

joined_data<-readRDS('../Data/joined_data.rds', refhook = NULL)
data_copy<-data_Wider
events_data<-na.omit(data_copy,cols=c("type_cross"))
```

## Showing all Data

### Short Data Variable(short_data)

<br/>
It contains the columns that represents th average of 15 minutes before and 15 minutes after each event occurs <br />
```{r}
head(short_Data)
```

### Bow Shock Data Variable (bs_data)
<br />
It contains only bow shock data. <br />
```{r}
kable(head(bs_data),"html")%>% scroll_box(width = "100%")
```

### Events Data Variable (events_data)

<br />
It contains only the data of all the events that occured (BS,MP,DG,SC) <br />
```{r}
kable(head(events_data),"html")%>% scroll_box(width = "100%")
```

## Analysing the Bowshock events and Magnetopause events from Dataset

### When cassini was in Outbound
This happened between the days 72 and 74 of the year 2005 <br />

#### Plotting Total Magnetic strength vs Time Stamp. 
 Also drawing a line that pointing the events.<br />
```{r fig.width=13}

###Finding all points in MAG 2005 between day 72 and 74

d1<-as.Date(72,origin="2005-01-01")
d2<-as.Date(73,origin="2005-01-01")
temp<-filter(joined_data,between(date,d1,d2))
bowShock<-filter(temp,type_cross=="BS")$TimeStamp
MagP<-filter(temp,type_cross=="MP")$TimeStamp

ggplot(data=temp, aes(x=TimeStamp , y=abs(BTotal.nT.))) +
  geom_line()+
  ylab("|B|(nT)")+
  xlab("Timestamp")+
  geom_vline(xintercept=bowShock, linetype="dotted",colour="blue",size=1.3)+
  geom_vline(xintercept=MagP, linetype="dotted",colour="red",size=1.3)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```
<br />
From the above figure the Red dotted line represents the time at which the Magnetopause region identified and blue dotted line represents the Bow Shock events.<br />


### When cassini was in Inbound
This happened between the days 136 and 138 of the year 2005 <br />

#### Plotting Total Magnetic strength vs Time Stamp. 
 Also drawing a line that pointing the events.<br />
```{r fig.width=10}

###Finding all points in MAG 2005 between day 72 and 74

d1<-as.Date(136,origin="2005-01-01")
d2<-as.Date(138,origin="2005-01-01")
temp<-filter(joined_data,between(date,d1,d2))
bowShock<-filter(temp,type_cross=="BS")$TimeStamp
MagP<-filter(temp,type_cross=="MP")$TimeStamp

ggplot(data=temp, aes(x=TimeStamp , y=abs(BTotal.nT.))) +
  geom_line()+
  ylab("|B|(nT)")+
  xlab("Timestamp")+
  geom_vline(xintercept=bowShock, linetype="dotted",colour="blue",size=1.3)+
  geom_vline(xintercept=MagP, linetype="dotted",colour="red",size=1.3)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```
<br />
From the above figure the Red dotted line represents the time at which the Magnetopause region identified and blue dotted line represents the Bow Shock events.<br />

## Plotting the trajectory of Cassini
<br/>
In this plot we are trying to identify the positions at which all the events occurred.
<br />
```{r fig.width=11, fig.height=7}

ggplot(data_Wider, aes(x=X_KSM.km. , y=Y_KSM.km.))+
  geom_point(alpha=0.3)+
  geom_point(events_data, mapping =aes(x=X_KSM.km. , y=Y_KSM.km.,color=type_cross,shape=dirn_cross))


```
<br/>
From this plot we can say that most of the events occurred in a particular region which faces the sun .
<br />


## Data Imbalance



### Proportion of Different directions in each event types.

<br/>
In this plot we are trying to identify the proportion of Total events in the dataset (large).
<br />
```{r fig.width=11}


ggplot(data = short_Data) +
  geom_bar(mapping = aes(x = type_cross, fill=dirn_cross))

```


### Proportion of each event types in Dataset.

<br/>
In this plot we are trying to identify the Count of each events from an events only dataset.
<br />
```{r fig.width=11}


ggplot(events_data, aes(type_cross, fill=dirn_cross )) +
  geom_bar(stat="count")

```


<br/>
From the plot we can clearly say that BowSchock events are recorded more times that Magneto pause.But  the proportion of Inward or Outward direction is same for both.
<br />


### Proportion of Different directions in each event types.

<br/>
In this plot we are trying to identify the proportion of each events in the dataset.
<br />
```{r fig.width=11}


ggplot(data = short_Data) +
  geom_bar(mapping = aes(x = type_cross, fill=dirn_cross) , position = "fill")

```